Lattice Boltzmann simulations of magnetically responsive\\ anisotropic particle stabilized emulsions

### 2025-03-06 chapters/final_remarks.tex
Changes 1. Added more analysis to application of magnetic fields during formation chapter 2. Added references to fix some (not all) literature reference issues 3. Added some detail in analysis techniques section
Changes 1. Added conclusion in final remarks 2. Added hysteresis curve for prolate particles stabilized bijels in aim2 section 3. Added description of synthesis techniques in lit review

@@

This work utilized 2 ellipsoidal particle geometries which were not as anisotropic as rods or plates that some particles have naturally, such as cellulose nanocrystals and graphene nanoplates. These particles have also been shown to naturally have nematic ordering at higher volume fractions that correspond to the volume fractions present at bijel interfaces. An investigation into how using rods or plate like particles would be instructive in identifying how onsager theory can be linked to bijel microstructure. \textcolor{blue}{https://doi.org/10.1039/D1SM00367D}

Colloidal systems made with cohesive and soft particles have been shown in the literature to have different glass transition points and rheological behavior from their hard sphere counterparts. In bijels, cohesive particles in particular have been suggested as a means to create "armored" bijels to improve their performance in catalytic materials, allowing higher flow rates to be used. Investigations into soft particles are of interest in biomedical engineering applications, with newly developed nanogels and hydrogels being suggested as drug carriers or vectors for stimuli response through temperature or pH.

@@

>>>>

Owing to their high surface area to volume ratio, porous materials are seeing a surge in popularity in applications
such as catalysis, battery electrodes and pharmaceuticals. Given their wide variety of applications, identifying 
fabrication techniques that allow access to the various pore length scales is of interest. One such synthesis technique
is emulsion templating which offers a wide variety of accessible microstructures, addition of stimuli response and 
large number of possible systems that can be fabricated. A microstructure that can be fabricated from emulsion templates
is the bicontinuous interfacially jammed emulsion gel (bijel).

Bijels are normally fabricated using Thermally Induced Phase Separation (TIPS). However, TIPS does not allow for continuous
fabrication of bijels. More recent fabrication techniques such as Solvent Transfer Induced Phase Separation (STrIPS) and 
Vapor Induced Phase Separation (VIPS) allow for continuous fabrication and access to various microstructures. Both techniques
require modifications to the initial emulsion mixture to facilitate microstructure adjustments, which affect the final rheological
properties of the system. Decoupling the microstructure and the casting mixture of the bijel would allow for greater flexibility
in the synthesis of the material.

Stimuli response has been used before to modify the microstructure of particle stabilized emulsions. Magnetic fields offer
targeted material response and low applied field strengths necessary for response. Past work investigating the effect of magnetic
stimuli on bijels stabilized with spherical particles yielded little microstructure change. However, anisotropic particles have
also been used to stabilize bijels. Anisotropic particles at interfaces under magnetic fields have been shown to tilt out of 
interfaces. These effects have yet to be captured in bijels stabilized by ellipsoidal particles under magnetic fields.
This study addresses this knowledge gap using a hybrid Molecular-Dynamics multicomponent method Lattice Boltzmann Method.
We split this work into three aims; Aim 1 addressed the microstructure obtained when applying a magnetic field of various strengths 
onto bijels stabilized by ellipsoidal particles, compared to bijels stabilized wih spherical particles. Aim 2 addressed the 
structural response of bijels stabilized by ellipsoidal particles and analyzes the effect of initial order of the particle monolayer
on the structural response observed. Aim 3 addressed the constant shear response of bijels stabilized by ellipsoidal particles
with pre-existing particle order and under magnetic fields.

In Aim 1, bijels stabilized with spherical particles do not respond to the application of a magnetic field. However, bijels stabilized
with ellipsoidal particles have an increase in the average domain size by 3 \%. The microstructure also becomes anisotropic, 
characterized as a change in the directional tortuosity and distribution of channel widths as a function of the distance from the 
interface. The curvature also becomes more negative as the applied field strength is increased. The microstructure anisotropy is caused
by direction specific jamming of the particle monolayer originating from particle ordering to the magnetic field.

When investigating the particle monolayer, the orientational order changes as a function of the applied field strength and has particle
morphology specific time evolution. The interface angle was also characterized and shown to have magnetic field dependence, although
this was attributed to the orientational ordering of the particles on the interface. When analyzing the ordering of particles on the
interface, disc-like particles see lowered local ordering as the magnetic field strength is increased while rod-like particles see
greater local ordering. These properties are attributed to how those particle morphologies prefer to orient themselves at interfaces
to one another, with disc like particles preferring to stack while rod like particles order side to side or end to end.

In Aim 2

In Aim 3

@@

This work utilized 2 ellipsoidal particle geometries chosen for comparisons to previous literature using this 
particle geometry. Particles based on cellulose nanocrystals or graphene nanoplates are now in use to fabricate
particle stabilized emulsions.These particles have also been shown to have capillary bridging and particle stacking,
not seen in the particles used in this work. These particles in bulk have been shown to have some intrinsic ordering
that can be predicted using onsager theory. An investigation into how using rods or plate like particles would be 
instructive in identifying if onsager theory can be used in bijel formation and its link to bijel microstructure. 
\textcolor{blue}{https://doi.org/10.1039/D1SM00367D}

Colloidal systems made with cohesive and soft particles have been shown in the literature to have different glass 
transition points and rheological behavior from their hard sphere counterparts. In bijels, cohesive particles in 
particular have been suggested as a means to create "armored" bijels to improve their performance in catalytic materials, 
allowing higher flow rates to be used. Investigations into soft particles are of interest in biomedical engineering 
applications, with newly developed nanogels and hydrogels being suggested as drug carriers or vectors for stimuli 
response through temperature or pH. LBM methods that implement the immersed boundary method can be used to model soft
particles with a DLVO potential used to model electrostatics between particles. \textcolor{blue}{https://doi.org/10.1039/D3SM01648J}

@@

% Extensions to this work can be accomplished by investigating the effect of applying a magnetic field on the bijel 
% while under shear. Ferrofluid models that predict bingham plastic like flows, $\frac{\eta}{eta_{f}} = 1 + \frac{Mn^{*}(\phi_p)}{Mn}$, 
% have been developed and defined using the Mason number, $Mn = \frac{8\eta_{f} \dot{\gamma}}{\mu_{0} \mu_{f} \beta^{2} H_0^2}$. 
% \textcolor{blue}{https://doi.org/10.1122/1.4935850, https://linkinghub.elsevier.com/retrieve/pii/S1359029405000385}

### 2025-03-06 chapters/lit_review.tex
Changes 1. Added conclusion in final remarks 2. Added hysteresis curve for prolate particles stabilized bijels in aim2 section 3. Added description of synthesis techniques in lit review

\textcolor{blue}{https://www.mdpi.com/2311-5521/5/3/150}

\section{Colloidal glasses}

\begin{itemize}
    \item https://pubs.acs.org/doi/10.1021/acsmacrolett.6b00826
    \item Add stuff on dynamic heterogeneity
    \begin{itemize}
        \item https://pubs.rsc.org/en/content/articlelanding/2012/sm/c2sm25267h
    \end{itemize}
    \item Add stuff on particle jamming
    \begin{itemize}
        \item https://journals.aps.org/rmp/pdf/10.1103/RevModPhys.82.2633
    \end{itemize}
    \item Add stuff on cooperatively rearranging regions
        \begin{itemize}
            \item https://doi.org/10.1038/ncomms4829
            \item https://journals.aps.org/prl/abstract/10.1103/PhysRevLett.110.188301
            \item https://journals.aps.org/prl/abstract/10.1103/PhysRevLett.107.065702
        \end{itemize}
    \item Add stuff on reentrant glass phenomena 
    \begin{itemize}
        \item https://doi.org/10.1209/0295-5075/86/58001
    \end{itemize}
\end{itemize}

>>>>

% \textcolor{blue}{https://www.mdpi.com/2311-5521/5/3/150}

% \section{Colloidal glasses}

% \begin{itemize}
%     \item https://pubs.acs.org/doi/10.1021/acsmacrolett.6b00826
%     \item Add stuff on dynamic heterogeneity
%     \begin{itemize}
%         \item https://pubs.rsc.org/en/content/articlelanding/2012/sm/c2sm25267h
%     \end{itemize}
%     \item Add stuff on particle jamming
%     \begin{itemize}
%         \item https://journals.aps.org/rmp/pdf/10.1103/RevModPhys.82.2633
%     \end{itemize}
%     \item Add stuff on cooperatively rearranging regions
%         \begin{itemize}
%             \item https://doi.org/10.1038/ncomms4829
%             \item https://journals.aps.org/prl/abstract/10.1103/PhysRevLett.110.188301
%             \item https://journals.aps.org/prl/abstract/10.1103/PhysRevLett.107.065702
%         \end{itemize}
%     \item Add stuff on reentrant glass phenomena 
%     \begin{itemize}
%         \item https://doi.org/10.1209/0295-5075/86/58001
%     \end{itemize}
% \end{itemize}

### 2025-03-06 chapters/paper2.tex
Changes 1. Added conclusion in final remarks 2. Added hysteresis curve for prolate particles stabilized bijels in aim2 section 3. Added description of synthesis techniques in lit review

Stimuli responsive emulsions have seen interest for use in fields such
as enhanced oil recovery, porous material templates, in-situ reaction
seperation systems and drug delivery scaffolds.
\cite{tang_stimuli-responsive_2016,} \cite{bago_rodriguez_capsules_2019}
These applications are enabled through stimuli response driven
microstructure modification, changing the permeability, sueface area and
morphology of the emulsion. Leveraging advances in particle synthesis
techniques and citing concerns over surfactant toxicity and
environmental concerns, particle stabilized emulsions have made a
resurgence in popularity after their discovery in 1907 by Ramsden and
Pickering. \cite{rozynek_opening_2019} \cite{zhou_magnetic_2011} Stimuli
response has been enabled in particle stabilized emulsions using pH,
temperature or chemical composition changes.
\cite{maingret_dextran-based_2020} \cite{lin_ph-responsive_2018}
\cite{raju_ph-responsive_2018} \cite{zhang_ph_2022}

Owing to its tortuous, co-continuous domain morphology, bicontinuous
interfacially jammed emulsion gels (bijel) are studied for their
possible applications to fabricate membranes, catalyst supports, battery
electrodes and pharmaceutical products. \cite{yabuno_preparation_2020}
\cite{samdani_bicontinuous_2017} \cite{cha_bicontinuous_2019}
\cite{garcia_scalable_2019} \cite{santiago_cordoba_aerobijels_2020}
Bijels are formed through the arresting of spinodal decomposition of
partially miscible fluids. \cite{stratford_colloidal_2005}
\cite{herzig_bicontinuous_2007} \cite{tavacoli_novel_2011} Once phase
separation begins, the incipient interface begins to sweep through the
system causing particles to begin to adsorbing into it. Once the
interfacial area matches that of the cross sectional area of the
adsorbed particles, the structure jams in place.
\cite{stratford_colloidal_2005} \cite{herzig_bicontinuous_2007}
\cite{tavacoli_novel_2011} After their discovery in simulations and
subsequent synthesis in experiments, a rich diversity of bijel
literature has developed around various fluid systems that allow for
bijel synthesis, methods to fabricate them at scale and their
applications.

@@

\section{Results}\label{sec:results_p2}
\subsection{Field strength dependence on domain
size}\label{field-strength-dependence-on-domain-size}

@@

\begin{figure} 
\centering 
\includegraphics[scale=0.4]{../figures/results/paper2/particle_viz-field_on.png} 
\caption{Visualizations of bijels stabilized with oblate and prolate particles simulated at $t = 0$ (left column) and $t = 10^5$ (right column). The first and third rows show the particle monolayer of bijels stabilized with oblate and prolate ellipsoids respectively with no applied field. The second and fourth rows show the particle monolayer of bijels stabilized with oblate and prolate ellipsoids respectively with an applied field of $\bar{B}_z = 1$. Particle reorientation to the direction of the field can be seen, resulting in coarsening of the domains.} 
\label{fig:particle_viz-field_on} 
\end{figure}

@@

\begin{figure} 
\centering 
\includegraphics[scale=0.35]{../figures/results/paper2/domain_size-field_on-scaled.png} 
\caption{Plots of domain size with different timescales. The top and bottom row represent the average domain size for bijels stabilized with oblate and prolate particles respectively. The figures on the left plots the domain size against time rescaled with the nematic to isotropic transition time, $\tau_S$. The figures in the middle plot the domain size against time rescaled with the interfacial capillary timescale, $\tau_{\langle \psi \rangle}$. The figures on the right plot the domain size against time rescaled with the particle rearrangement timescale, $\tau_{\langle Q_6 \rangle}$. We show that the controlling timescale when applying a magnetic field onto bijels with disordered particles is the latter timescale, indicating that the dominating effect is steric hindrance of the particles during rearrangment at the interface.} 
\label{fig:domain_size-field_on-scaled} 
\end{figure}

@@

\begin{figure} 
\centering 
\includegraphics[scale=0.4]{../figures/results/paper2/particle_viz-field_up.png} 
\caption{Visualizations of bijels stabilized with oblate and prolate particles simulated at $t = 0$ (left column) and $t = 10^5$ (right column). The first and third rows show the particle monolayer of bijels stabilized with oblate and prolate ellipsoids respectively with an applied field of $\bar{B}_z = 0.2$. The second and fourth rows show the particle monolayer of bijels stabilized with oblate and prolate ellipsoids respectively with an applied field of $\bar{B}_z = 1$. We increase the applied field to $\bar{B}_z = 1$ in both cases and show that particle reorientation to the field occurs at the $\bar{B}_z = 0.2$ but not in the $\bar{B}_z = 1$ case.} 
\label{fig:particle_viz-field_up} 
\end{figure}

@@

\begin{figure} 
\centering 
\includegraphics[scale=0.5]{../figures/results/paper2/interface_angle-nint-field_up.png} 
\caption{Plots of the average interfacial angle $\psi$ and the number of particles on the interface $n_{int}$ on the left and right respectively, plotted against time for bijels stabilized with ellipsoidal particles with aspect ratio 2. When looking at $\psi$, a change in dynamics can be seen as the initial particle order increases. When looking at $n_{int}$, different dynamics are seen above and below the isotropic to nematic transition point. Both results suggest that the ordering of particles is linked to the dynamics of domain coarsening.} 
\label{fig:interface_angle-nint-field_up} 
\end{figure}

@@

\begin{figure} 
\centering 
\includegraphics[scale=0.5]{../figures/results/paper2/interface_angle-nint-field_down.png} 
\caption{Plots of the average interfacial angle $\psi$ and the number of particles on the interface $n_{int}$ on the left and right respectively, plotted against time for bijels stabilized with ellipsoidal particles with aspect ratio 2. When looking at $\psi$, the $\bar{B}:1 \rightarrow 1$ case looks distinctly different from the other processes. At $\bar{B} = 1$, the particles are totally ordered to the field. This suggests that the presence of some particle disorder resists domain coarsening.} 
\label{fig:interface_angle-field_down} 
\end{figure}

>>>>

Stimuli-responsive emulsions have attracted significant interest for applications in enhanced oil recovery, porous material templating, 
in-situ reaction separation systems, and drug delivery scaffolds \cite{tang_stimuli-responsive_2016, bago_rodriguez_capsules_2019}. 
These applications are facilitated by stimuli-induced microstructural modifications, which alter the permeability, surface area, and 
morphology of the emulsion. Advancements in particle synthesis techniques, along with growing concerns over surfactant toxicity and 
environmental impact, have contributed to the renewed interest in particle-stabilized emulsions since their initial discovery by Ramsden 
and Pickering in 1907 \cite{rozynek_opening_2019, zhou_magnetic_2011}. Stimuli-responsive behavior in these emulsions has been achieved 
through changes in pH, temperature, or chemical composition 
\cite{maingret_dextran-based_2020, lin_ph-responsive_2018, raju_ph-responsive_2018, zhang_ph_2022}.

Due to their tortuous, co-continuous domain morphology, bicontinuous interfacially jammed emulsion gels (bijels) have been 
extensively studied for applications in membrane fabrication, catalyst supports, battery electrodes, and pharmaceutical 
products \cite{yabuno_preparation_2020, samdani_bicontinuous_2017, cha_bicontinuous_2019, garcia_scalable_2019, santiago_cordoba_aerobijels_2020}.  
Bijels form through the arrested spinodal decomposition of partially miscible fluids 
\cite{stratford_colloidal_2005, herzig_bicontinuous_2007, tavacoli_novel_2011}. During phase separation, the emerging interface propagates 
through the system, driving particle adsorption. Once the available interfacial area matches the total cross-sectional area of the adsorbed 
particles, the structure becomes jammed in place \cite{stratford_colloidal_2005, herzig_bicontinuous_2007, tavacoli_novel_2011}. Since their 
initial discovery in simulations and subsequent experimental synthesis, a diverse body of literature has emerged, exploring fluid systems 
compatible with bijel formation, scalable fabrication techniques, and potential applications.

@@

\section{Results}\label{sec:results_p2}
\subsection{Hysteresis curve}\label{section:hysteresis_curve}

Bijels have proposed applications as soft matter based material templates, drug delivery vectors and separations systems. The materials 
performance in each application is dependent upon the microstructure of the bijel template. Many of the applications listed above benefit 
from stimuli response, allowing for in-situ microstructure tuning to modify permeability of the material. We want to understand whether 
bijels provide access to reversible structural modifications which would allow stimuli responsive bijels to be used in in-situ applications.

We begin our investigation by characterizing the microstructure changes observed on bijels stabilized with prolate ellipsoids with a 
particle volume fraction of $\phi_p = 0.1$. We apply a constant magnetic field of $\bar{B}_z = 0.2$ onto bijels that were simulated 
with no magnetic field. Once the microstructure has reached its new state, we ramp up the magnetic field to $\bar{B}_z = 0.5$ and 
finally $\bar{B}_z = 1$. We then apply the opposite procedure to reduce the applied field strength, resulting in a magnetic field 
ramp of $\bar{B}_z: 0 \rightarrow 0.2 \rightarrow 0.5 \rightarrow 1 \rightarrow 0.5 \rightarrow 0.2 \rightarrow 0$.We characterize the 
microstructure changes observed through calculating a characteristic length scale, called the average domain size. We calculate the hysteresis 
curve by calculating the domain size at the final timestep of each magnetic field change. We then plot these domain sizes against the magnetic field 
applied to obtain them and split them in the descending and ascending magnetic field directions to demonstrate the differences observed. We plot 
these results in Figure \ref{fig:hysteresis_curve}

\begin{figure} 
    \centering 
    \includegraphics[scale=0.5]{../figures/results/paper2/hysteresis_curve.png} 
    \caption{Plot of the hysteresis curve of a bijel stabilized by magnetically responsive prolate particles. We observe that the domain size increases as we increase the applied
    magnetic field strength. However, upon decrease of the field strength, we see that the microstructure does not return to its previous value.} 
    \label{fig:hysteresis_curve} 
\end{figure}

Figure \ref{fig:hysteresis_curve} illustrates the hysteresis behavior of magnetically responsive bijels. We observe a hysteresis response of the 
bijel microstructure, characterized by a non-return of the microstructure to the original average domain size value when reducing the applied 
magnetic field strength. The increase in the average length scale when applying the magnetic field is due to domain coarsening caused by 
reorientation of particles to the magnetic field strength, reducing the interfacial coverage. When decreasing the magnetic field strength, 
we observe that the domain size does not return to the original value meaning that that the microstructure of the bijel is unaffected by 
the reduction in magnetic field. This suggests that the particles at the interface are in a kinetically trapped configuration.

\subsection{Field strength dependence on domain
size}\label{section:field-strength-dependence-on-domain-size}

@@

\begin{figure} 
\centering 
\includegraphics[scale=0.2]{../figures/results/paper2/particle_viz-field_on.png} 
\caption{Visualizations of bijels stabilized with oblate and prolate particles simulated at $t = 0$ (left column) and $t = 10^5$ (right column). The first and third rows show the particle monolayer of bijels stabilized with oblate and prolate ellipsoids respectively with no applied field. The second and fourth rows show the particle monolayer of bijels stabilized with oblate and prolate ellipsoids respectively with an applied field of $\bar{B}_z = 1$. Particle reorientation to the direction of the field can be seen, resulting in coarsening of the domains.} 
\label{fig:particle_viz-field_on} 
\end{figure}

@@

\begin{figure} 
\centering 
\includegraphics[scale=0.4]{../figures/results/paper2/domain_size-field_on-scaled.png} 
\caption{Plots of domain size with different timescales. The top and bottom row represent the average domain size for bijels stabilized with oblate and prolate particles respectively. The figures on the left plots the domain size against time rescaled with the nematic to isotropic transition time, $\tau_S$. The figures in the middle plot the domain size against time rescaled with the interfacial capillary timescale, $\tau_{\langle \psi \rangle}$. The figures on the right plot the domain size against time rescaled with the particle rearrangement timescale, $\tau_{\langle Q_6 \rangle}$. We show that the controlling timescale when applying a magnetic field onto bijels with disordered particles is the latter timescale, indicating that the dominating effect is steric hindrance of the particles during rearrangment at the interface.} 
\label{fig:domain_size-field_on-scaled} 
\end{figure}

@@

\begin{figure} 
\centering 
\includegraphics[scale=0.2]{../figures/results/paper2/particle_viz-field_up.png} 
\caption{Visualizations of bijels stabilized with oblate and prolate particles simulated at $t = 0$ (left column) and $t = 10^5$ (right column). The first and third rows show the particle monolayer of bijels stabilized with oblate and prolate ellipsoids respectively with an applied field of $\bar{B}_z = 0.2$. The second and fourth rows show the particle monolayer of bijels stabilized with oblate and prolate ellipsoids respectively with an applied field of $\bar{B}_z = 1$. We increase the applied field to $\bar{B}_z = 1$ in both cases and show that particle reorientation to the field occurs at the $\bar{B}_z = 0.2$ but not in the $\bar{B}_z = 1$ case.} 
\label{fig:particle_viz-field_up} 
\end{figure}

@@

\begin{figure} 
\centering 
\includegraphics[scale=0.4]{../figures/results/paper2/interface_angle-nint-field_up.png} 
\caption{Plots of the average interfacial angle $\psi$ and the number of particles on the interface $n_{int}$ on the left and right respectively, plotted against time for bijels stabilized with ellipsoidal particles with aspect ratio 2. When looking at $\psi$, a change in dynamics can be seen as the initial particle order increases. When looking at $n_{int}$, different dynamics are seen above and below the isotropic to nematic transition point. Both results suggest that the ordering of particles is linked to the dynamics of domain coarsening.} 
\label{fig:interface_angle-nint-field_up} 
\end{figure}

@@

\begin{figure} 
\centering 
\includegraphics[scale=0.4]{../figures/results/paper2/interface_angle-nint-field_down.png} 
\caption{Plots of the average interfacial angle $\psi$ and the number of particles on the interface $n_{int}$ on the left and right respectively, plotted against time for bijels stabilized with ellipsoidal particles with aspect ratio 2. When looking at $\psi$, the $\bar{B}:1 \rightarrow 1$ case looks distinctly different from the other processes. At $\bar{B} = 1$, the particles are totally ordered to the field. This suggests that the presence of some particle disorder resists domain coarsening.} 
\label{fig:interface_angle-field_down} 
\end{figure}
